---
title: "Universal_Bank_Using_Naive_Bayes_Model"
output:
  html_document: default
  pdf_document: default
date: "2024-03-06"
---

# Executive Summary-

#The exercise conducted a thorough examination of the correlation between specific customer behaviours and the likelihood of accepting personal loan offers from Universal Bank.

#A Comparative Analysis- was conducted to evaluate the probabilities generated by the Naive Bayes classifier in contrast to those manually calculated utilising the pivot tables. The respective values of 11.056% and 11.065% indicate a substantial level of similarity. This finding signifies that the model-based approach and the empirical data are consistent, implying that the Naive Bayes classifier offers an appropriate prediction for this particular situation. . 

# Step 1: Importing Relevant Datasets in R

```{r}
library(caret)
```
#Loading Caret dataset for data partitioning

# Step 2: Data Analysis and Data Prep

```{r}
Universal_bank<- read.csv("~/Downloads/UniversalBank.csv")
head(Universal_bank)
str(Universal_bank)
```
#head() function is used here to display the first few rows.

#str() function is used here to display the structure of the data.


# Step 3: Dividing the Data into Testing and Training

```{r}
set.seed(123)
UB_Train<- createDataPartition(Universal_bank$Personal.Loan,p=0.6,list=FALSE)
Training_data<-Universal_bank[UB_Train,]
Validation_data<-Universal_bank[-UB_Train,]
```

# Step 4: Creating a Pivot Table

#First, we need to convert the data

```{r}
Training_data$Online<-as.factor(Training_data$Online)
Training_data$CreditCard<-as.factor(Training_data$CreditCard)
Training_data$Personal.Loan<-as.factor(Training_data$Personal.Loan)
```
#Before creating the pivot table we need to first convert the data into categorical data.

#Creating the Pivot Table After Converting The Data

```{r}
pivot_table <- xtabs(~ CreditCard + Personal.Loan + Online, data = Training_data)
print(pivot_table)
View(pivot_table)
```

#Values from the Pivot table

```{r}
num_loan_accepts_cc_online = 57 # CC = 1, Online = 1, Loan = 1
total_cc_online = 57+475 # CC = 1, Online = 1 (both Loan = 1 and Loan = 0)
```

#From the table, we have:

#Number of customers with CC = 1, Online = 1, and Loan = 1, 57

#Total number of customers with CC = 1 and Online = 1 (regardless of Loan status): 57 (Loan = 1) + 475 (Loan = 0)


#Calculating the Probability

```{r}
probability_acceptance_cc_online = num_loan_accepts_cc_online / total_cc_online
print(probability_acceptance_cc_online)
```

#The probability that a customer will accept the personal loan offer, given that they have a bank credit card (CC = 1) and are an active user of online banking services (Online = 1), is approximately 10.71%.


# Step 5: Creating Pivot Table with Loan as Rows and Online as Columns

```{r}
pivot_Online<-xtabs(~ Personal.Loan + Online, data = Training_data)
print(pivot_Online)
```

# Step 6: Creating Pivot Table with Loan as Rows and Credit Card as Columns

```{r}
pivot_cc <- xtabs(~ Personal.Loan + CreditCard, data = Training_data)
print(pivot_cc)
```

# Step 7: Probabilities Calculation From The Above-Mentioned Pivot Tables 

```{r}

# P(CC = 1 | Loan = 1)
p_cc_given_loan1 <- 91 / (91 + 187)

# P(Online=1|Loan=1)
p_online_given_loan1 <- 179 / (99 + 179)

# P(Loan = 1)
p_loan1 <- (91 + 187) / (1930 + 792 + 91 + 187)

# P(CC=1|Loan=0)
p_cc_given_loan0 <- 792 / (1930 + 792)

# P(Online=1|Loan=0)
p_online_given_loan0 <- 1620 / (1102 + 1620)

# P(Loan = 0)
p_loan0 <- (1930 + 792) / (1930 + 792 + 91 + 187)

# Output the probabilities
probabilities <- list(
  p_cc_given_loan1 = p_cc_given_loan1,
  p_online_given_loan1 = p_online_given_loan1,
  p_loan1 = p_loan1,
  p_cc_given_loan0 = p_cc_given_loan0,
  p_online_given_loan0 = p_online_given_loan0,
  p_loan0 = p_loan0
)
 print(probabilities)
```

#The calculated probabilities based on the data from the provided screenshots are as follows:

#P(CC=1∣Loan=1) - The probability of having a credit card given that the loan is accepted: 32.73%

#P(Online=1∣Loan=1) - The probability of being an online user given that the loan is accepted: 64.39%

#P(Loan=1) - The probability of loan acceptance: 9.27%

#P(CC=1∣Loan=0) - The probability of having a credit card given that the loan is not accepted: 29.10%

#P(Online=1∣Loan=0) - The probability of being an online user given that the loan is not accepted: 59.52%

#P(Loan=0) - The probability of loan non-acceptance: 90.73%

#Calculating the Overall Probabilities

```{r}
P_cc <- p_cc_given_loan1 * p_loan1 + p_cc_given_loan0 * p_loan0
print(P_cc)
P_online <- p_online_given_loan1 * p_loan1 + p_online_given_loan0 * p_loan0
print(P_online)
```

#Overall probability of a customer having a credit card regardless of whether they accept a loan or not is approximately 29.43%.

#Overall probability of a customer using online services, irrespective of their loan acceptance status is approximately 59.97%.

#Calculating the Naive Bayes Probability

```{r}
p_loan1_given_cc1_online1 <- (p_cc_given_loan1 * p_online_given_loan1 * p_loan1) / (P_cc * P_online)

print(p_loan1_given_cc1_online1)
```

#The naive-bayes probability that a customer will accept the personal loan offer, given that they have a bank credit card (CC = 1) and are an active user of online banking services (Online = 1), is approximately 11.065%.

# Question F- Compare this value with the one obtained from the pivot table in (B). Which is a more accurate estimate?

#Answer- Empirical Probability: The empirical probability (10.71%) is an accurate reflection of the historical data as it is derived from the actual occurrence of events in the training data.The resulting representation accurately captures the true characteristics of the dataset and remains unaffected by any model assumptions.

#Naive Bayes Model-Based Probability: The Naive Bayes model-based probability (11.065%) represents the estimated probability obtained from the model constructed using the training data. An estimate derived from a probabilistic model is the naive Bayes probability. 

#The empirical probability provides the most accurate approximation of the training data observations. In this particular instance, 10.71%


# Step 8: Creating Naive Bayes Model

```{r}
library(e1071)
nb_model <- naiveBayes(Personal.Loan ~ CreditCard + Online, data = Training_data) 
```

#Creating a New Data Frame For The Customer With CC = 1 and Online = 1

```{r}
new_customer <- data.frame(CreditCard = factor(1, levels = levels(Training_data$CreditCard)),
                           Online = factor(1, levels = levels(Training_data$Online)))
```

#Getting The Predicted Probabilities For The New Customer Data

```{r}
predicted_probs <- predict(nb_model, new_customer, type = "raw")
print(predicted_probs)
```

#The Probability of Loan Acceptance For a Customer With CC = 1 and Online = 1

```{r}
loan_acceptance_prob <- predicted_probs[1, "1"]
print(loan_acceptance_prob)
```

#Naive Bayes model's estimated probability that a customer with a credit card (CC = 1) and who is an active user of online banking services (Online = 1) will accept the personal loan offer (Loan = 1) is approximately 11.056%

# Question G -Compare this to the number you obtained in (E).

#11.065% calculated using numbers from your pivot tables.

#11.056% obtained by running the Naive Bayes algorithm in R.

#Answer-The two probabilities are very close to each other, which indicates that the model's assumptions,(people with credit card and who are actively using online services will accept personal loan) well for this particular subset of the data.

#Both methods are just about equally good at predicting the chance of someone taking a personal loan under these conditions.This similarity is a positive sign for the validity of using the Naive Bayes model in this context.

